Classifying galaxy spectra at $0.5<z<1$: an unsupervised approach

### 2017-06-24 answer_to_second_referee_comments.tex

We thank the reviewer for his/her helpful comments and apologize for the delay in returning a revised version of the manuscript. Large blocks of changed text in the manuscript appear in bold (we have also made minor grammatical and style changes which are not marked). Our responses to individual comments are in the text below, beginning with ---

@@

This paper proposes the use of self-organizing maps (1D and 2D) to aid in classifying galaxy spectra.  The revision has addressed several points and improve the clarity of the text.  However, there are several remaining issues.

1. The authors did not address my comment about the significance of the work.  What was learned?  What method is advocated, and what is the evidence for its superiority?

The sentence that was added about 1D vs. 2D maps comes very late in the paper and is quite abstract, stating that the two methods are complementary.  The Introduction should establish the goals of the investigation so the reader understands the purpose of each experiment as well as its significance.

--- We have reorganized the last two paragraphs of Section 1 and added text to better describe the goal of the investigation: comparing different methods of classifying galaxy spectra against templates, with a focus on self-organizing maps.
We have added additional text to Sections 4.2 and 5 comparing 1D and 2D SOMs.
We have also added some text to Section 5 discussing the significance of the work.

The paper claims that "Compared to previous attempts to classify this sample, the self-organizing map approach was able to better classify the spectra which had similarities with more than one template." (abstract), but I cannot find a quantitative comparison with any previous results.  Figure 9 is presented as the result to compare with previous work, but since the previous results are not presented nor summarized, the reader can't tell how they differ.  The paper reports "more precision", but this is not defined; does it mean smaller error bars?  Stronger correlation?  How does this show "better classification"?  To support such a claim, this paper needs to select a performance metric and report on the values achieved in these experiments, compared to other methods.
--- We selected several performance metrics. Two of them (an agreement score and the Fleiss kappa score) measure the degree to which the classifications agree. 
% TBD: have not done the following yet
As a measure of the classification performance, we calculate the dispersion in properties of the galaxies in each class and compare these values between methods. The results of this analysis are discussed in Section 4.4.

@@

I had suggested comparing the results to a simple nearest-neighbor classification baseline, but this was not addressed.
--- We apologize for missing this helpful suggestion in the previous version. This is now part of the analysis described in the paragraph above; the results appear in Section 4.4.

The comment about how SOMs provide an ordered clustering of the data (with respect to feature values), which k-means does not, is certainly true.  But the paper must then explain why this is important or valuable.
--- We have added text to the end of both Section 4.2 (formerly section 4.1) and Section 5 explaining the utility of an ordered clustering.

2. Perhaps I have missed it somewhere, but I don't think the paper ever states how exactly the classification of new data is done.  As I understand it, the authors train an SOM using template spectra, then classify new spectra by determining which SOM node is the best fit for the new item and classifying it based on the labeled template spectra that also reside in that node -- but this isn't actually stated anywhere.  Please add text explaining the classification step.  This would probably go in the final paragraph of the introduction and/or a new subsection in section 3 ("Method") - this section describes SOMs and k-means but not how to use either model to classify new data. (i.e., the SOM can't classify anything without knowing the labels of its original data set.  So although it builds the SOM in an unsupervised way, it requires the supervision of the template labels to be able to assign a label to a new item).
-- Added the following paragraph  at the end of section 3.2: 
    "As mentioned in the introduction, one of the main advantages of the SOM method  is that the resulting networks can be used to cluster new datasets with no additional training required. 
     New data with the same dimensionality as the original input data can be presented to the already trained network: the SOM algorithm finds the best matching unit considering the weight of nodes in the trained network and each vector from the new data set.
     The winner node determines the place of each new vector on the SOM.
     The location of the new vectors on the SOM allows them to be compared with the original data set."

@@

On page 2, lines 29-31, the authors added this text: "An SOM can be trained without supervision and its results used to classify other data. Therefore, this method can be considered as (semi)-supervised." As written, the phrase "this method" seems to refer to SOMs (so the statement wouldn't be true - SOMs are fully unsupervised), but I think instead the authors' proposed method (of building an SOM and then applying it to classify new data, using labels on the SOM-grouped data) is what is meant.  Please clarify.
--- Changed to: 
    "The training of a self-organizing map is fully unsupervised.
     Using the resulting map as a template to classify other data requires labelling the output, which is why some groups consider self-organizing maps to be a (semi)-supervised method."

3. The inclusion of k-means helps give useful context to this work. Figure 10 shows a head-to-head comparison of k-means and SOM applied to the 12 templates using 4 clusters/nodes.  But I can't tell which one is better.  The text states that they are similar but "show discrepancies for spectra types between the extreme cases".  Which ones?  How can I tell from this figure?
--- The text describing the differences between the K-means and SOM classification of the templates (in particular, which templates are grouped together by the two methods) has been rewritten and moved to Section 4.1 (newly added in this revision).

At the bottom of page 10, the paper states: "The other difference between these two methods is that the network created by the SOM method can be used to classify the other spectra, while for the K-means clusters they can only be used for one set of data."  This isn't true.  (The paper itself points out that SOM reduces to k-means if the weight on the neighborhood function goes to 0, so you should be able to use the result in the same way.)  You can (and should, to do a direct comparison) use the clusters just like you used the SOM nodes - for a new data point, assign it to the most similar cluster center, then assign the template label from the original data used to construct the clusters.
--- We thank the referee for this helpful suggestion and have implemented it in our use of K-means. The text has been changed to remove this incorrect statement.

To analyze the 142 galaxies, the SOM and k-means experiments used different methodology and so are difficult to compare or understand. It seems that for k-means, the authors clustered the 142 sample galaxies and then classified the mean spectrum in each cluster and reported how well that did.  But for the SOM, the authors constructed the SOM based on the 12 templates from K96 and then assigned classes to individual new spectra (not the nodes).  This doesn't seem comparable.  Please use the same methodology in both cases.
--- This is a reasonable criticism and we have changed our methodology to match between methods. We now use the templates to generate the initial classification in both cases, and then apply that classification to the 142 sample galaxies.
%TBD: Sahar please check.

Further, k-means and the SOM used different distance functions (Euclidean vs. spectral angle distance) when operating on the same data.  Differences in results could simply be due to this change.  The authors' explanation for this choice does not make sense to me.
--- We have changed to using the Euclidean distance in both cases.
%TBD: Sahar please check.

The text says that Figure 11 shows "the best and the worst fitting results among the 22 clusters" (for k-means) but instead it seems to show the best-fit for two clusters (9 and 22) which both look pretty good.  The text implies that Fig 11 (right) is a poor match, but it looks very good.  Please explain.
--- We apologize for the confusion. As we have substantially changed the comparison between K-means and SOM results, this figure is no longer relevant and has been removed.

### 2017-06-24 main.tex
PB edits

\begin{abstract}
    The spectrum of a galaxy contains information about its physical properties.
    Classifying spectra using templates helps elucidate the nature of a galaxy's energy sources.
    However, available templates are limited, which can complicate spectral classification.
    Most current classification methods classify spectra outside the range of templates with low accuracy, and will fail in special cases.
    In this paper, we introduce a new method of classifying the spectra
    of high redshift galaxies: the self-organizing map.
    Self-organizing maps are a type of artificial neural network, mostly used to identify subgroups in a sample.
    We trained semi-supervised self-organizing map networks using a set of
    templates covering the wavelength range from far ultraviolet to near infrared. 
    The trained networks were used to classify the spectra of a sample of 142 galaxies with $0.5 < z < 1$.
    Compared to previous attempts to classify this sample, the
    self-organizing map approach was able to better classify the spectra which had similarities with more than one template.
    The median spectra and physical properties of the galaxies in each neuron of the self-organizing maps are consistent with other results on galaxy evolution.

@@

    Self-organizing maps can be used to classify celestial objects (e.g. stars, quasars, spectra of galaxies, light curves, etc.)
    Galaxy spectra can be classified in various networks with different sizes/dimensions based on the information needed from the data. 
    If a broad and general classification is required, networks can have one dimension with a few neurons. 
    If one needs more detailed classifications, a higher number of neurons should be used.
    Since self-organizing maps do not include the uncertainty of input parameters, sometimes too much attention to detail can cause problems in classifications. 
    When using the SOM method, one should consider whether small differences between objects are physically meaningful when separating two groups from each other.

    We used SOMs to classify the template spectra of \citetalias{Kinney96}, made from galaxies with known morphological type, and created networks with different uses.
    By varying the size of the networks, we found the relative similarity between the \citetalias{Kinney96} template classes.
    Two-dimensional networks can be used to find the relations of the neurons and classifications in great detail.
    The sample galaxies have more freedom to be categorized, and only 23 of them occupy exactly the same neurons as the \citetalias{Kinney96} template spectra (in one-dimensional networks this number was 56).
    Additionally, by using two-dimensional networks we showed that we do not have any outliers in the sample galaxies.

    A one-dimensional network with 22 neurons was needed in order to
    separate all 12 \citetalias{Kinney96} spectra; we concluded that \citetalias{Kinney96} types B and E, and types SB1 and SB2, are very similar to each other.
    We also showed that networks generated with the SOM method can be used to easily identify new types of spectra in large surveys.

    A sample of 142 high-redshift galaxy spectra from \citetalias{Hossein12} was classified by the trained networks.
    The classification results showed that using SOMs can allow identification of galaxies with spectra similar to two or more morphological types.
    The freedom of having in-between types is one of the main differences between supervised and unsupervised artificial neural networks.
    \citetalias{Hossein12} used a supervised training method to train networks with \citetalias{Kinney96} spectral templates and classify the same sample of 142 galaxy spectra;
    however, this method could not classify 37 out of 142 spectra.
    The unsupervised artificial neural network method used in this project was able to classify  all 142 spectra in the sample
    as belonging to one of the morphological types introduced by \citetalias{Kinney96} or a class in between those types.
    \textbf{Although it might seem that the classes of spectra using SOMs are arbitrary, we remind the reader that this classification emerges from the results of data mining on the previous templates.}

    The properties of the galaxies in the \citetalias{Hossein12} sample using the new classification were found to be more tightly correlated in mean values of age, specific star formation rate, stellar mass, and far-UV extinction than in previous studies. 
    The properties of the galaxies in each group are in good agreement with their morphological types.
    This particular sample of high-redshift galaxies is well-described by the range of the \citetalias{Kinney96} templates; some of the spectra fall in between template classes. 
    This implies that these galaxies are not markedly different from those found in the local universe.
    The self-organizing map method can order the spectra in such a way as to also match the order along the age/SFR and SSFR/M* relation.
     Other classification methods such as K-means and supervised artificial neural networks can provide a general classification with similar performance to the SOM. The ordered results and intermediate classes produced by self-organizing maps are distinct advantages of this method which can be utilized for other astronomical applications.
%TBD: add some text here explainng why ordering is important

\section*{ACKNOWLEDGMENTS}
The authors thank the referee for insightful comments which helped us to improve the manuscript.
The authors thank S. Lianou, A. Tammour, S.C. Gallagher, R.G. Abraham, M. Daley and A. Sigut for their useful comments. 
S.R. and P.B. also acknowledge research support from the Natural Sciences and Engineering Research Council of Canada. 
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%biblio
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\bibliographystyle{apalike}
\bibliography{ref_mining_h.bib}

>>>>

\begin{abstract}
    The spectrum of a galaxy contains information about its physical properties.
    Classifying spectra using templates helps elucidate the nature of a galaxy's energy sources.
    In this paper, we introduce a new method of classifying the spectra
    of galaxies: the self-organizing map.
    Self-organizing maps are a type of artificial neural network, mostly used to identify subgroups in a sample.
    We trained semi-supervised self-organizing map networks using a set of
    templates covering the wavelength range from far ultraviolet to near infrared. 
    The trained networks were used to classify the spectra of a sample of 142 galaxies with $0.5 < z < 1$.
    %following sentences still need work.
    Compared to previous attempts to classify this sample, the
    self-organizing map approach was able to better classify the spectra which had similarities with more than one template.
    The median spectra and physical properties of the galaxies in each neuron of the self-organizing maps are consistent with other results on galaxy evolution.

@@

    Self-organizing maps can be used to classify celestial objects (e.g. stars, quasars, spectra of galaxies, light curves, etc.).
    {\bf In this work we present a detailed comparison between different types of self-organizing maps and other methods used to classify galaxy spectra.
    Based on our experience we suggest some general guidelines for use of SOM in galaxy spectral classification.}
    If a broad and general classification is required, networks can have one dimension with a few neurons. 
    If one needs more detailed classifications, a higher number of neurons should be used.
    Since self-organizing maps do not include the uncertainty of input parameters, sometimes too much attention to detail can cause problems in classifications. 
    When using the SOM method, one should consider whether small differences between objects are physically meaningful when separating two groups from each other.

    We used SOMs to classify the template spectra of \citetalias{Kinney96}, made from galaxies with known morphological type, and created networks with different uses.
    By varying the size of the networks, we found the relative similarity between the \citetalias{Kinney96} template classes, which can be roughly ordered in one dimension by their amount of star formation.
     A one-dimensional network with 22 neurons was needed in order to
    separate all 12 \citetalias{Kinney96} spectra; we concluded that \citetalias{Kinney96} types B and E, and types SB1 and SB2, are very similar to each other.
    \textbf{Two-dimensional networks allow more freedom in the galaxy spectral classification.
    Training $12\times 12$ neurons with the \citetalias{Kinney96} spectra, we found that the two dimensions of the resulting self-organized maps corresponded roughly to amounts of star formation and extinction.}

    In the two-dimensional network, only 23 galaxies occupied exactly the same neurons as the \citetalias{Kinney96} template spectra (in one-dimensional networks this number was 56).
    Additionally, by using two-dimensional networks we showed that there are no outliers in the sample galaxies.

    A sample of 142 high-redshift galaxy spectra from \citetalias{Hossein12} was classified by the trained networks.
    The classification results showed that using SOMs can allow identification of galaxies with spectra similar to two or more morphological types.
    The freedom of having in-between types is one of the main differences between supervised and unsupervised artificial neural networks.
    \citetalias{Hossein12} used a supervised training method to train networks with \citetalias{Kinney96} spectral templates and classify the same sample of 142 galaxy spectra;
    however, this method could not classify 37 out of 142 spectra.
    The self-organizing map (unsupervised artificial neural network) method used in the present work was able to classify  all 142 spectra in the sample
    as belonging to one of the morphological types introduced by \citetalias{Kinney96} or a class in between those types.
    \textbf{Although it might seem that the classes of spectra using SOMs are arbitrary, we remind the reader that this classification emerges from the results of data mining on the previous templates.
    Compared to other classification methods, we found that self-organizing maps show similar levels of agreement with a baseline chi-squared template matching method.}

    The properties of the galaxies in the \citetalias{Hossein12} sample using the new classification were found to be more tightly correlated in mean values of age, specific star formation rate, stellar mass, and far-UV extinction than in previous studies. 
    % QUANTIFY IN ABOVE SENTENCE. BELOW SENTENCE: DO WE ACTUALLY SHOW THIS? 
    The properties of the galaxies in each group are in good agreement with their morphological types.
    This particular sample of high-redshift galaxies is well-described by the range of the \citetalias{Kinney96} templates although some of the spectra fall in between template classes.
    This implies that these galaxies are not markedly different from those found in the local universe.
    The self-organizing map method can order the spectra in such a way as to also match the order along the age/SFR and SSFR/M* relations.
    Such an ordered classification does not naturally arise from other classification methods such as K-means and supervised artificial neural networks, and can be used to investigate underlying physical causes of changes in properties between classification groups.

\section*{ACKNOWLEDGMENTS}
The authors thank the referee for insightful comments which helped us to improve the work.
The authors thank S. Lianou, A. Tammour, S.C. Gallagher, R.G. Abraham, M. Daley and A. Sigut for their useful comments. 
S.R. and P.B. also acknowledge research support from the Natural Sciences and Engineering Research Council of Canada. 
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%biblio
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\bibliographystyle{apalike}
\bibliography{ref_mining_h.bib}

### 2017-06-24 sections0.08/data0.08.tex
PB edits

\section{Data}
\label{sec: data_highZ}
We use spectral templates from \citetalias{Kinney96} (in 10~\AA\ wavelength bins) to train neural networks. 
We classify the spectra of 142 galaxies at $0.5<z<1$ from \citetalias{Hossein12} using the trained networks, and use their physical properties to test the new classifications.
Following the \citetalias{Hossein12} work, we chose these two sets of data not only to show the application of SOMs in spectral clustering, but also to easily compare supervised and unsupervised methods.

 \subsection{Kinney Spectral Templates}
     \begin{figure}
        \centering
        \includegraphics[width=0.5\textwidth]{images/k96.jpg}
        \caption[\citet{Kinney96} spectral templates for 12 types of galaxies]{\citet{Kinney96} spectral templates for 12 types of galaxies, from fig. 1 in \citet{Hossein12}. The type of each template is shown in each frame. Plots B, E, S0, Sa, Sb and Sc show spectra that belong to quiescent galaxies. Starburst galaxy spectra are indicated with SB1 to SB6. Higher numbers represent more intrinsic extinction. Note that the vertical axis label from \citet{Hossein12} is incorrect:  the quantity plotted in the vertical axis is flux density, not flux.}
        \label{fig: k96}
    \end{figure}

@@

 \subsection{SED and Properties of the Sample Galaxies} 
    \citetalias{Hossein12} selected 142 galaxies from the spectroscopic campaign of the ESO GOODS-South field~\citep{Vanzella05, Vanzella06, Vanzella08}.
    The 142 galaxies were selected based on the availability of photometry from HST/ACS, VLA/ISAAC, and {\it Spitzer}/MIPS and IRAC (10--13 filters with $\sim 0.4<\lambda<24~\mu$m in the observed frame).
   Data from these instruments was necessary in order to have a complete picture of stellar population and star formation rate. 
    For each galaxy, a robust spectroscopic redshift and photometric measurements from the GOODS-MUSIC catalogue \citep{Santini09} were available.
%  We have gone back and forth on the following sentence several times. The commented-out version (what you had) doesn't make sense to me. I replaced it with what I understand was done, based on reading T12.
%   \citetalias{Hossein12} matched the point spread function (PSF) of the photometric data with \citetalias{Kinney96} data 
\citetalias{Hossein12} used point-spread-function matched photometry 
    from \citet{Santini09} as input to the Code Investigating GALaxy Emission ({\em CIGALE});~\citep[][hereafter N09]{Noll09} to generate the best-fit SED for each galaxy.
    \citetalias{Hossein12} produced the best SED match, with the wavelength interval of 910~\AA~to $\sim 80$~cm, for each galaxy\footnote{This wavelength interval is the default output of the {\em CIGALE} code.}.
    Assuming decreasing SFR and visual attenuation ($\tau$) model, Salpeter initial mass function~\citep{Salpeter55}, and an old stellar population with age of $\sim 10$~Gyr, he derived physical properties of the galaxies such as age and stellar mass.
    Some of these properties are shown in Table~\ref{tab: props}.
    In Section~\ref{sec: 1D_somz}, we study these properties for each category.
    More details on creating SEDs and extracting information about galaxy properties using {\em CIGALE} can be found in \citetalias{Noll09} and \citetalias{Hossein12}.

@@

    We classify SEDs that were produced by \citetalias{Hossein12} using the created networks.
    These SEDs are publicly available~\footnote{All the SEDs produced by \citetalias{Hossein12}  can be found at this \href{http://telbib.eso.org/detail.php?bibcode=2012AJ....144..172T}{ESO webpage}.} in the form of flux per rest frame wavelength over a wide range of wavelengths.
    Since we have chosen the \citetalias{Hossein12} SEDs to be classified by the trained network, we only used the part of the SEDs that have the same wavelength range as the \citetalias{Kinney96} spectral templates.

>>>>

\section{DATA}
\label{sec: data_highZ}
We use spectral templates from \citetalias{Kinney96} (in 10~\AA\ wavelength bins) to train neural networks. 
We classify the spectra of 142 galaxies at $0.5<z<1$ from \citetalias{Hossein12} using the trained networks, and use their physical properties to test the new classifications.
Following the \citetalias{Hossein12} work, we chose these two sets of data not only to show the application of SOMs in spectral clustering, but also to easily compare supervised and unsupervised methods.

 \subsection{Kinney spectral templates}
     \begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{images/k96.jpg}
        \caption {\citet{Kinney96} spectral templates for 12 types of galaxies, from fig. 1 in \citet{Hossein12}. The type of each template is shown in each frame. Plots B, E, S0, Sa, Sb and Sc show spectra that belong to quiescent galaxies. Starburst galaxy spectra are indicated with SB1 to SB6. Higher numbers represent more intrinsic extinction. Note that the vertical axis label from \citet{Hossein12} is incorrect:  the quantity plotted in the vertical axis is flux density, not flux.}
        \label{fig: k96}
    \end{figure}

@@

 \subsection{SED and Properties of the Sample Galaxies} 
    \citetalias{Hossein12} selected 142 galaxies from the spectroscopic campaign of the ESO GOODS-South field~\citep{Vanzella05, Vanzella06, Vanzella08}.
    The 142 galaxies were selected based on the availability of photometry from HST/ACS, VLA/ISAAC, and {\it Spitzer}/MIPS and IRAC (10--13 filters with $\sim 0.4<\lambda<24~\mu$m in the observed frame).
   Data from these instruments was necessary in order to have a complete picture of stellar population and star formation rate. 
    For each galaxy, a robust spectroscopic redshift and photometric measurements from the GOODS-MUSIC catalogue \citep{Santini09} were available.
\citetalias{Hossein12} used point-spread-function matched photometry 
    from \citet{Santini09} as input to the Code Investigating GALaxy Emission ({\em CIGALE});~\citep[][hereafter N09]{Noll09} to generate the best-fit SED for each galaxy.
    \citetalias{Hossein12} produced the best SED match, with the wavelength interval of 910~\AA~to $\sim 80$~cm, for each galaxy\footnote{This wavelength interval is the default output of the {\em CIGALE} code.}.
    Assuming decreasing SFR and visual attenuation ($\tau$) model, Salpeter initial mass function~\citep{Salpeter55}, and an old stellar population with age of $\sim 10$~Gyr, he derived physical properties of the galaxies such as age and stellar mass.
    Some of these properties are shown in Table~\ref{tab: props}.
    In Section~\ref{sec: 1D_somz}, we study these properties for each category.
    More details on creating SEDs and extracting information about galaxy properties using {\em CIGALE} can be found in \citetalias{Noll09} and \citetalias{Hossein12}.

@@

    We classify SEDs that were produced by \citetalias{Hossein12} using the created networks.
    These SEDs are publicly available\footnote{All the SEDs produced by \citetalias{Hossein12}  can be found at this \href{http://telbib.eso.org/detail.php?bibcode=2012AJ....144..172T}{ESO webpage}.} in the form of flux per rest frame wavelength over a wide range of wavelengths.
    Since we have chosen the \citetalias{Hossein12} SEDs to be classified by the trained network, we only used the part of the SEDs that have the same wavelength range as the \citetalias{Kinney96} spectral templates.

### 2017-06-24 sections0.08/intro0.08.tex
PB edits

\section{Introduction}
\label{sec: intro_somz}
%General information about SEDs
Nearly all information we can obtain from a galaxy is encapsulated in the light it emits; every observable phenomenon in a galaxy leaves a footprint on its spectral energy distribution (SED).
We can determine some physical properties of a galaxy by properly modeling various features observed in its SED, and the general shape of the SED can be used as an identifier of the morphological type.
Considering the main features of SEDs, galaxies can be categorized into two main groups: quiescent or star-forming.
Each group has its own characteristic features and can in turn be divided into many sub-branches.

%%Creating templates and NIR_UV templates
Several attempts have been made to create a complete set of templates for categorizing the spectral type of galaxies using observations of nearby galaxies~(e.g.~\citealt{Kinney93}, \citealt[][hereafter \citetalias{Kinney96}]{Kinney96}, \citealt{Bershady00}, \citealt{Mannucci01}). 
Based on their usage, these templates are restricted to certain wavelengths.
In the near-infrared (NIR) to ultraviolet (UV) wavelengths (the region where the energy output of stars peaks), the spectrum\footnote{In the remainder of this paper, we are concerned with a limited range in wavelength, and so we refer to spectra rather than spectral energy distributions.} contains information about the main physical properties of galaxies, e.g., age, star formation rate (SFR), stellar mass, metallicity, and interstellar medium. 
\textbf{TEXT HERE RE: WHY WE CARE ABOUT CLASSIFYING
(1) do galaxy spectra form a continuous distribution or can they be separated into discrete groups as represented by the \citetalias{Kinney96} templates? (2) if the distribution is continuous, how do galaxy properties vary along the spectral distribution?
}

% categorizing galaxies based on their SED

@@

%ANNs
Artificial neural networks (ANNs), which are inspired by the way neurons in a human brain route and process data, are very powerful tools that are used in data processing and pattern recognition problems.
An ANN contains many interconnected units (nodes or neurons) which process data and work together to solve problems.
It uses a set of training methods to learn about nonlinear and complex relations between input and output data, and how to apply these relations to new sets of data~\citep[e.g.][]{Hossein14,Hossein16a,Hossein16b,Ellison16a, Ellison16b}.
Studies have shown that ANNs outperform chi-square minimizing techniques and can be used as an alternative choice for fitting data~\citep[e.g.][]{Marquez91}.
Specifically, ANNs perform faster in large databases~\citep[][]{Gulati97}.

%Training methods for networks - some grammar edits 20170122
Neural networks can be trained using two methods: supervised and unsupervised.
In supervised methods, a neural network is trained using input data based on a desired outcome.
These methods are very useful for classification of data with specific target values.
In unsupervised methods there is no prediction of output:
these methods classify data based on their underlying structures and hidden patterns.
Unsupervised methods are very helpful in obtaining knowledge from the data, or when the underlying structure of the data is not well established.

%SOM
A Kohonen self-organizing map (also called self-organizing map, or SOM) is a (semi)-supervised neural network for mapping and visualizing a complex and nonlinear high dimension data introduced by~\citet{Kohonen82}.
It shows simple geometrical relationships in non-linear high dimensional data on a map \citep{Kohonen98}.
\textbf{The training of a self-organizing map is fully unsupervised.
     Using the resulting map as a template to classify other data requires labelling the output, which is why some groups consider self-organizing maps to be a (semi)-supervised method.}

%SOM in Astronomy
The utilization of the SOM in astronomy dates back to the 1990s, with \citet[][]{Odewahn92}, \citet[][]{Hernandez94}, and \citet[][]{Murtagh95} among the first to use SOMs in their studies.
\citet{Geach12} used COSMOS data to demonstrate two of the main applications of SOMs: object classification and clustering, and photometric redshift estimation. The latter has been the subject of many other studies \citep[e.g.][]{Kind14a}.
From classifying quasars' spectra to star/galaxy classifications, from gamma-ray burst clustering to classification of light curves, this method has proved to be useful in various fields of astronomy \citep[e.g.][]{Maehoenen95, Miller96, Andreon00, Balastegui01, Rajaniemi02, Brett04, Scaringi09}.

%SOM in Astronomy continued.
Large spectroscopic surveys have made available integrated spectra of millions of galaxies.
These integrated spectra combine the light of billions of individual stars and nebulae within a galaxy, and
finding patterns and common characteristics between galaxies can be a complex task.
\citet{In12} introduced a new clustering tool based on the SOM method for analyzing these large datasets.
They used $\sim 60000$ spectra from the Sloan Digital Sky Survey \citep[SDSS;][]{Abazajian09} to test their tool, and created very large SOMs to analyze the type of spectra/objects.
They also generated SOMs from quasars' spectra in order to find unusual types of spectra. 
Later, \citet{Meusinger16} used these SOMs and updated data from SDSS and other surveys to find a new class of quasars.
The other application of SOMs is to find outliers or errors in the data.
\citet{Fustes13} produced a package based on SOM to classify spectra from the GAIA survey that were previously classified as ``unknown'' by the SDSS pipeline. This package can distinguish an astronomical object from instrumental artifact, and then classify the object based on its spectrum.

@@

%20170122: combined and reorganized last 2 paragraphs of section - DRAFT
%%%%After second referee report
\textbf{
The goal of this work is to compare different methods of classifying galaxy spectra against templates, with a particular focus on self-organizing maps. 
The questions we want to answer are: 
(1) What is the degree of agreement between $\chi^2$ fitting, one- and two-dimensional self-organizing maps, K-means clustering and artificial neural networks? 
(2) Are certain spectral types more (or less) consistently classified by the different methods? 
(3) To what extent do the average properties of galaxy classes represent a continuous distribution of physical properties?
We analyze the same sample of 142 galaxies with $0.5 < z < 1$ as 
\citet[][hereafter \citetalias{Hossein12}]{Hossein12}, who classified the spectra using a supervised (multi-layer feed-forward) neural network method with the \citetalias{Kinney96} templates.
For each of the 142 galaxies, we find the best chi-squared fit among the 12 templates  to have a base for comparing the methods' performance.
We also use the Fleiss Kappa index \citet{Fleiss71} to test interrater reliability  
between the three methods.
}

%not sure if we want this sentence or not
%Our hypothesis was that since a self-organizing map has the freedom to classify objects in between known classes, it could be applied to these data to find the best spectral class for the 37 galaxies that could not be classified by the supervised method.
% also not sure if we want these sentences.
% As in \citetalias{Hossein12}, we examine the averaged properties of the new galaxy groups and compare them with previous work.
%\citetalias{Hossein12} also 
%showed that there are tight correlations between physical properties of galaxies, and these %correlations might be different for each type of galaxy.
 %HERE IS WHERE WE MENTION OUR METRIC TO SAY WHICH METHOD IS BETTER

>>>>

\section{Introduction}
\label{sec: intro_somz}

Nearly all information we can obtain from a galaxy is encapsulated in the light it emits; every observable phenomenon in a galaxy leaves a footprint on its spectral energy distribution (SED).
We can determine some physical properties of a galaxy by properly modeling various features observed in its SED, and the general shape of the SED can be used as an identifier of the morphological type.
Considering the main features of SEDs, galaxies can be categorized into two main groups: quiescent or star-forming.
Each group has its own characteristic features and can in turn be divided into many sub-branches.

Several attempts have been made to create a complete set of templates for categorizing the spectral type of galaxies using observations of nearby galaxies~(e.g.~\citealt{Kinney93}, \citealt[][hereafter \citetalias{Kinney96}]{Kinney96}, \citealt{Bershady00}, \citealt{Mannucci01}). 
Based on their usage, these templates are restricted to certain wavelengths.
In the near-infrared (NIR) to ultraviolet (UV) wavelengths (the region where the energy output of stars peaks), the spectrum\footnote{In the remainder of this paper, we are concerned with a limited range in wavelength, and so we refer to spectra rather than spectral energy distributions.} contains information about the main physical properties of galaxies, e.g., age, star formation rate (SFR), stellar mass, metallicity, and interstellar medium. 
\textbf{
Using templates to classify galaxy spectra can help to answer the question of whether galaxy spectra form a continuous distribution or can  be separated into discrete groups.
Templates are also helpful in computing $K$-corrections and discovering new classes of objects as outliers \citep{Folkes96}.
}

@@

Artificial neural networks (ANNs), which are inspired by the way neurons in a human brain route and process data, are very powerful tools that are used in data processing and pattern recognition problems.
An ANN contains many interconnected units (nodes or neurons) which process data and work together to solve problems.
It uses a set of training methods to learn about nonlinear and complex relations between input and output data, and how to apply these relations to new sets of data~\citep[e.g.][]{Hossein14,Hossein16a,Hossein16b,Ellison16a, Ellison16b}.
Studies have shown that ANNs outperform chi-square minimizing techniques and can be used as an alternative choice for fitting data~\citep[e.g.][]{Marquez91}.
Specifically, ANNs perform faster in large databases~\citep[][]{Gulati97}.

Neural networks can be trained using two methods: supervised and unsupervised.
In supervised methods, a neural network is trained using input data based on a desired outcome.
These methods are very useful for classification of data with specific target values.
In unsupervised methods there is no prediction of output:
these methods classify data based on their underlying structures and hidden patterns.
Unsupervised methods are very helpful in obtaining knowledge from the data, or when the underlying structure of the data is not well established.

A Kohonen self-organizing map (also called self-organizing map, or SOM) is a (semi)-supervised neural network for mapping and visualizing a complex and nonlinear high dimension data introduced by~\citet{Kohonen82}.
It shows simple geometrical relationships in non-linear high dimensional data on a map \citep{Kohonen98}.
\textbf{The training of a self-organizing map is fully unsupervised.
     Using the resulting map as a template to classify other data requires labelling the output, which is why some groups consider self-organizing maps to be a (semi)-supervised method.}

The utilization of the SOM in astronomy dates back to the 1990s, with \citet[][]{Odewahn92}, \citet[][]{Hernandez94}, and \citet[][]{Murtagh95} among the first to use SOMs in their studies.
\citet{Geach12} used COSMOS data to demonstrate two of the main applications of SOMs: object classification and clustering, and photometric redshift estimation. The latter has been the subject of many other studies \citep[e.g.][]{Kind14a}.
From classifying quasars' spectra to star/galaxy classifications, from gamma-ray burst clustering to classification of light curves, this method has proved to be useful in various fields of astronomy \citep[e.g.][]{Maehoenen95, Miller96, Andreon00, Balastegui01, Rajaniemi02, Brett04, Scaringi09}.

Large spectroscopic surveys have made available integrated spectra of millions of galaxies.
These integrated spectra combine the light of billions of individual stars and nebulae within a galaxy, and
finding patterns and common characteristics between galaxies can be a complex task.
\citet{In12} introduced a new clustering tool based on the SOM method for analyzing these large datasets.
They used $\sim 60000$ spectra from the Sloan Digital Sky Survey \citep[SDSS;][]{Abazajian09} to test their tool, and created very large SOMs to analyze the type of spectra/objects.
They also generated SOMs from quasars' spectra in order to find unusual types of spectra. 
Later, \citet{Meusinger16} used these SOMs and updated data from SDSS and other surveys to find a new class of quasars.
The other application of SOMs is to find outliers or errors in the data.
\citet{Fustes13} produced a package based on SOM to classify spectra from the GAIA survey that were previously classified as ``unknown'' by the SDSS pipeline. This package can distinguish an astronomical object from instrumental artifact, and then classify the object based on its spectrum.

@@

\textbf{
The goal of this work is to investigate the use of self-organizing maps in classifying galaxy spectra against templates, and compare them to other methods. 
The questions we want to answer are: 
(1) What is the degree of agreement between $\chi^2$ fitting, one- and two-dimensional self-organizing maps, K-means clustering and artificial neural networks? 
(2) Are certain spectral types more (or less) consistently classified by the different methods? 
(3) To what extent do the average properties of galaxy classes represent a continuous distribution of physical properties?
We analyze the same sample of 142 galaxies with $0.5 < z < 1$ as 
\citet[][hereafter \citetalias{Hossein12}]{Hossein12}, who classified the spectra using a supervised (multi-layer feed-forward) neural network method with the \citetalias{Kinney96} templates.
For each of the 142 galaxies, we find the best chi-squared fit among the 12 templates  to have a base for comparing the methods' performance.
We also use the Fleiss Kappa index \citet{Fleiss71} to test interrater reliability 
between the three methods.
% HAVE NOT YET DONE THE FOLLOWING. MAY WANT TO INCLUDE COMMENTED SENTENCES.
We calculate the dispersion in properties between galaxies classified in the same groups in order to measure how well each classification method groups similar objects together.
}

% As in \citetalias{Hossein12}, we examine the averaged properties of the new galaxy groups and compare them with previous work.
%\citetalias{Hossein12} also 
%showed that there are tight correlations between physical properties of galaxies, and these %correlations might be different for each type of galaxy.
